New lower bounds for Schur and weak Schur numbers

### 2021-07-05 article/article.tex
Nouveau style biblio
ws 6

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. We also generalized the concept 
of "template" to weak Schur numbers. Finding new templates allows us to get explicit partitions improving lower bounds 
as well as the growth rate for Schur numbers, weak Schur numbers and Ramsey numbers \(R_n(3)\). The new lower 
bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS (9) \geqslant 22\,536\) and 
\(\WS (10) \geqslant 71\,214 \).

@@

\qquad We would first like to thank our professors, Joanna Tomasik and Arpad Rimmel, who guided and advised
us during the whole research and writing. Last but not least, we would like to address warm thanks
to Fred Rowley who kindly answered our many questions and with who it was a pleasure to exchange.

\bibliography{biblio}
\bibliographystyle{IEEEtran}

>>>>

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. We also generalized the concept 
of "template" to weak Schur numbers. Finding new templates allows us to get explicit partitions improving lower bounds 
as well as the growth rate for Schur numbers, weak Schur numbers and Ramsey numbers \(R_n(3)\). The new lower 
bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS(6) \geqslant 646\), \(\WS (9) \geqslant 22\,536\) and 
\(\WS (10) \geqslant 71\,214 \).

@@

\qquad We would first like to thank our professors, Joanna Tomasik and Arpad Rimmel, who guided and advised
us during the whole research and writing. We also would like to address warm thanks
to Fred Rowley who kindly answered our many questions and with who it was a pleasure to exchange. 
Last but not least, we would like to thank Marijn Heule for providing the backdoors used in the computation of the 
new lower bound for \(\WS(6)\).

\bibliography{biblio}
\bibliographystyle{elsarticle-num.bst}

### 2021-07-05 article/weakschur.tex
ws 6

\qquad In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers. By analogy with the previous section,
we then introduce WS-templates as well as an associated sequence \(\WS^+(n)\). We find templates and use them to
provide new lower bounds for weak Schur numbers.

@@

\begin{corollary}
Let \(n,k \in \mathbb{N}^*\) and set \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then :
\[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}.
\]
\end{corollary}

@@

>>>>

\qquad In this section, we generalize Rowley's constructions for weak Schur numbers \cite{RowleyWS} and give an analogous
for weak Schur numbers of Abbott and Hanson's construction for Schur numbers. By analogy with the previous section,
we then introduce WS-templates as well as an associated sequence \(\WS^+(n)\). We find templates and use them to
provide new lower bounds for weak Schur numbers. Finally, we give a short explanation for the new lower bound 
\(\WS(6) \geqslant 646\) which was not directly obtained with a template contrary to the other lower bounds given in this 
paper.

@@

\begin{corollary}
\label{Corollary3.19}
Let \(n,k \in \mathbb{N}^*\) and set \( b_{max} = \max \{b\in \mathbb{N}^*/ \WS_b^+(n+1) = \WS^+(n+1)\}\).
Then :
\[
\WS(n+k) \geqslant S(k) \WS^+(n+1)+b_{max}.
\]
\end{corollary}

@@

\subsection{\(\WS(6) \geqslant 646\)}
\qquad When applying the inequality \(\WS(n+1) \geqslant 4 \, S(n) + 2\), one may realize that it is possible to build a weakly 
sum-free partition of length \(4 \, S(n) + 3\) for small values of \(n\) (\(n \leqslant 4\)) by using the contruction of 
\hyperref[Corollary3.19]{Corollary 3.19} for the integers 1, 2, \(4i\) and \(4i+1\) for \(i \in [\![1, S(n)]\!]\) but not 
constraining the other integers. We did the same for the Schur partitions corresponding to Schur number five. More precisely, 
we imposed these constraints only for \(i \leqslant 50\) in order to have more degrees of freedom. The number 50 was chosen 
arbitrarly so that all of the Schur number five partitions could tested in a few hours.

However, trying out all of the 2\,447\,113\,088 Schur number five partitions (\cite{Heule2017}) one by one would not result in a 
reasonable computation time: it is necessary to test the construction on several partitions at once. We encoded the problem as a 
satisfiability problem and used the 1616 backdoors that were used in \cite{Heule2017} in order to encode a group of partitions in 
a compact and efficient way. Among all of the 1616 backdoors, only the 911\textsuperscript{th} backdoor gave a weakly sum-free 
partition of length 643. This backdoor gave a weakly sum-free partition of length 646 and cannot give a weakly sum-free 
partition of length 647. A weakly sum-free partition of length 646 can be found in the \hyperref[WS(6)]{appendix}.
